MLH1 Results Outline

### 2020-02-25 doc/Results.rmd
starting on evo framework section

@@

G.HETC.val #1.2
LEW.HETC.val  #1
WSB.HETC.val  #1
KAZ.HETC.val  #1.1

SKIVE.HETC.val # 0.96
PWD.HETC.val  #0.88
MSM.HETC.val #0.9
#kable(head(main.table),  caption = "Table 1. MLH1 counts per cell summary statistics by strain")

@@

#sum( (main.table  %>% filter(sex == "male"))$Ncells)
#the kable version is bulkier than the other versaion
```

CALCULATE THE MAGNITUDES of HETC

We used counts of MLH1 foci per cell to estimate the genome wide recombination rates across our panel of `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild derived inbred strains sampled from three subspecies within the house mouse species complex;  _M. m. domesticus,_ _M. m. musculus_. and _M. m. molossinus_. MLH1 foci were quantified from a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells)` spermatocytes and `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells)` oocytes. <!-- state the average number of mice-->
The sampling of prophase oocytes is diffecult and was the limiting factor thus there are some strains with only MLH1 counts from spermatocytes, which are included for sex specific analysises.

<!--
1. Fig1A  magnitude and direction of heterochiasmy;
-general patterns i) female higher and ii) low degree of hetC
(all strain means are above the expected min)

2. Fig1B and C mouse means and sex specific patterns; 
female distributions more random / uniform
male distributions are more clustered, low and high

@@

<!-- transition from first part into female -->

<!-- female section -->

@@

<to better understand this varation in RR within an evolutionary framework>

In order to test the effects of subspecies, sex and strain on the mean of MLH1 counts per cell, we fit `r length(DF.HetC.MixedModel.HQ$mean_co )` mouse averages of MLH1 foci per cell to a series of mixed and linear models.

<!--
(*note on sampling /sample numbers, the numbers of mice across strains, is shown in supplemental table )

**M1.Mixed Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex + rand(strain) + \varepsilon $$

Our model, M1., was designed to test the effects of the subspecies, sex, their interaction, and strain as a random effect on the mouse means.

@@

All of the effects within the mixed model were significant (LRT: subsp p= `r MLH1.MMlmer_results$sub.sp`, sex p = `r MLH1.MMlmer_results$sex.results`, subspecies * sex p = `r MLH1.MMlmer_results$interaction.pval`) and random LTR; strain p= `r MLH1.MMlmer_results$rand.pval`).

@@

@@

@@

```{r HetC.rates, echo=FALSE, include=FALSE, warning=FALSE}
#calculate the initial HetC levels from best quality mice and cells (compare both levels (cell and mouse average levels))
#subset MLH1_DF, remove non passing mice (as diagnoised from plots)
#remove quality 5 cells

```{r HetC.plot, echo=FALSE, warning=FALSE}
#HETC plot

#set up data

Het.C.plot
#this plot is missing several

>>>>

```{r MLH1.HetC.rates, echo=FALSE, include=FALSE, warning=FALSE}
#calculate the initial HetC levels from best quality mice and cells (compare both levels (cell and mouse average levels))
#subset MLH1_DF, remove non passing mice (as diagnoised from plots)
#remove quality 5 cells

```{r MLH1.HetC.plot, echo=FALSE, warning=FALSE}
#HETC plot
#change the main data?

@@

#G.HETC.val #1.2
#LEW.HETC.val  #1
#WSB.HETC.val  #1
#KAZ.HETC.val  #1.1

#SKIVE.HETC.val # 0.96
#PWD.HETC.val  #0.88
#MSM.HETC.val #0.9

#kable(head(main.table),  caption = "Table 1. MLH1 counts per cell summary statistics by strain")

@@

#sum( (main.table  %>% filter(sex == "male"))$Ncells)
#the kable version is bulkier than the other versaion

We used counts of MLH1 foci per cell to estimate the genome wide recombination rates across our panel of `r length(unique(  (main.table  %>% filter(species == "M.musculus") )$strain) )` wild derived inbred strains sampled from three subspecies within the house mouse species complex;  _M. m. domesticus,_ _M. m. musculus_. and _M. m. molossinus_. MLH1 foci were quantified from a total of `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "male"))$Ncells)` spermatocytes and `r sum( (main.table %>% filter(species == "M.musculus") %>% filter(sex == "female"))$Ncells)` oocytes. <!-- state the average number of mice
The sampling of prophase oocytes is diffecult and was the limiting factor thus there are some strains with only MLH1 counts from spermatocytes, which are included for sex specific analysises. -->

<!--
1. Fig1A  magnitude and direction of heterochiasmy;
-general patterns i) female higher and ii) low degree of hetC  (range of 1 to 1.2) (human is 1.6)
-All strain means are above the expected minimum of 20 (1CO per chromosome)
-there are exceptions they are followed up in the sex-specific plots  (male higher)

2. Fig1B and C mouse means and sex specific patterns; 
-female mouse mean distributions are have more spread across the range of means
-male distributions are more clustered, low and high (~30% evolution largest-rapid evolution of gwRR)
-sex-specific evolution, both sexes don't show equivalent elevation of gwRR
-Dom vs Molossinus and Musculus

@@

Figure 1A shows the strain averages for each sex together - this is a way to display the general patterns of sexual dimorphism (or heterochiasmy) across strains. 
Taking note of the direction and the magnitude our results confirm two general patterns that i) the genome wide recombination rates averages are greater females compared to males and ii) the degree of heterochiasmy (F:M rates) is generally low, rangeing from `r G.HETC.val` in _G_ to `r LEW.HETC.val` in _LEW_.

The notable exceptions to this pattern are from strains with male averages that exceed the female averages, with heterochiasmy values of `r PWD.HETC.val`, `r MSM.HETC.val` and `r SKIVE.HETC.val` for PWD, MSM and SKIVE respectively.

Examining the mouse means for the genome wide recombinaiton rate seperately for each sex reveals disctinct patterns of variation (Figure1 B and C).

The female means can be discribed as more evenly distributed around the sex-wide mean of ~25

While in males, the strain specific means -- can be seperated into groups of high genome wide recombination rates (near 30 COs per cell) and low genome wide recombiantion rates, near 23 COs per cell.

<!-- transition from first part into female -->
<!-- female section -->

@@

<to better understand this varation in RR within an evolutionary framework. Make table of models>

**M1. Mixed Model** $$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex + rand(strain) + \varepsilon $$

We applied a series of models to our dataset of mean MLH1 foci counts from `r length(DF.HetC.MixedModel.HQ$mean_co )` mice to fit the patterns of variation across sex, subspecies and strain into a evolutionary framework.

@@

All of the effects within the mixed model M1. were highly significant (LRT: subsp p= `r MLH1.MMlmer_results$sub.sp`, sex p = `r MLH1.MMlmer_results$sex.results`, subspecies * sex p = `r MLH1.MMlmer_results$interaction.pval`) and random LTR; strain p= `r MLH1.MMlmer_results$rand.pval`). Confirming the qualitative observations of wide spread variance. <!-- indicating XXX -->

@@

The linear models (M2. and M3.) allow more == specific comparisons

@@

<!--
In order to test the effects of subspecies, sex and strain on the mean of MLH1 counts per cell, we fit `r length(DF.HetC.MixedModel.HQ$mean_co )` mouse averages of MLH1 foci per cell to a series of mixed and linear models.

(*note on sampling /sample numbers, the numbers of mice across strains, is shown in supplemental table )

@@
